Lower Bounds on the Minimax Risk for the \\ Source Localization Problem

### 2017-02-05 paper.tex
Final version of the paper, without appendix.

We give lower bounds on the minimax risk (mean squared-error in location) in
estimating the location of the source, which apply to all estimators, for
certain classes of diffusive media, when using a uniformly distributed sensor
array. We show that for sensors of a fixed size, the lower bound decays with
increasing numbers of sensors. We also perform a preliminary numerical analysis
to understand the effect of increasing the number of sensors to infinity by
shrinking their size, wherein the bound saturates for large sensor numbers.  In
the second scenario, it is seen that there is greater benefit to increasing the
number of sensors as the signal-to-noise ratio increases.  Our bounds are the
first to give a scaling for the minimax risk in terms of the number of sensors
used.

@@

Hence, we believe that this paper is the first to give lower bounds on the
minimax error in estimating the location of a point source, which is observed
by sensors through a diffusive medium, and corrupted by additive white Gaussian
noise. We are also the first to analyze the implications physical sensor models
on such lower bounds. A simplistic theoretical model might assume that the
number of sensors can increase unbounded, while maintaining a fixed SNR at each
sensor. We show that this is not physically possible in our setting, and derive
numerical bounds which have radically different asymptotic behaviour.

@@

In the ``integrator sensor'' model, sensors have a \emph{width} which decreases
as their number grows to infinity. Concretely, the \mbox{$k$-th} sensor which
is located at $s = kS/m$ \emph{occupies} the space $((k{-}\frac{1}{2})S/m,
(k{+}\frac{1}{2})S/m)$.  The signal sensed by this sensor is given by the
$k$-th element of $\v x(\theta)$,
\begin{equation}
	x_k(\theta) = \int_{(k-\frac{1}{2})S/m}^{(k+\frac{1}{2})S/m} x(s;\theta) ds.
\end{equation}
Here, a white noise process $\epsilon(s)$ additively corrupts the
continuous-space signal $x(s;\theta)$ prior to sampling (see
Fig.~\ref{fig:signal-model}). $\v \epsilon$ is generated by integrating
$\epsilon(s)$ over a sensor width. This leads to i.i.d.\ noise values at
sensors (since sensors have no overlap), with variance inversely proportional
to $m$, i.e., $\sigma(m) \propto 1/\sqrt{m}$ (refer
Appendix~\ref{app:integrator-snr} in the full version of this
paper~\cite{FullVersion} for an explanation).

@@

	section~\ref{sec:source-localization}, the minimax risk in estimating the
	location of a point source is lower bounded by

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=2.8in]{lb-vs-m-annotated}
	\caption{A numerically generated plot of the lower bound on minimax risk of
		source localization, against the number of uniformly distributed
		sensors, $m$, used to estimate the location (code available
		online~\cite{Code}). The different curves are for different numbers of
		``trials'', $n$ (number of i.i.d.\ sensor observations). The bound
		saturates for large $m$, indicating that for a \emph{fixed} number of
		trials, there is decreasing benefit to using more sensors.  However,
	with more trials, the bound saturates at increasingly larger numbers of
sensors. Thus, there may be benefit to increasing the number of sensors when
you have more trials.}
	\label{fig:numerical}
\end{figure}

Plotting this numerically-computed bound for different numbers of trials
reveals some interesting trends (see Fig.~\ref{fig:numerical}). First, the
bound saturates as $m$ grows large, matching the behaviour of the
information-theoretic bound. Second, the lower bound on error is larger for
smaller numbers of sensors, as we expect. Third and most importantly, as the
number of trials increases, the bound begins to saturate only at a larger
number of sensors. In other words, there is \emph{increased} benefit in using
larger numbers of sensors, when we \emph{also} have more trials. This
substantiates the intuition given in~\cite{Grover2016Information}, which makes
this point using a Nyquist-rate argument: a larger number of sensors will
sample a larger number of high-frequency components of the source signal (and
hence allow us to locate it better), but only if the noise floor is
sufficiently low. Noise variance, in turn, can be reduced by averaging over
multiple trials.

@@

We also note that the minimax bounds from compressive sensing
literature~\cite{AriasCastro2013Fundamental} \emph{do not} apply in our
setting, because those consider the problem of recovering the \emph{whole}
source signal, $x(s;\theta)$. The loss function used there is hence the energy
of the difference signal, $\int \abs{x(s) - \hat x(s)}^2 ds$. We, on the other
hand, are interested in bounding the minimax error in \emph{location}, i.e.,
$\abs{\theta - \hat\theta}^2$.

>>>>

We give lower bounds on the minimax risk (mean squared-error in location) in
estimating the location of the source, which apply to all estimators, for
certain classes of diffusive media, when using a uniformly distributed sensor
array. We show that for sensors of a fixed size, the lower bound decays with
increasing numbers of sensors. We also analyze a more physical sensor model to
understand the effect of shrinking the size of sensors as their number
increases to infinity, wherein the bound saturates for large sensor numbers.
In this scenario, it is seen that there is greater benefit to increasing the
number of sensors as the signal-to-noise ratio increases.  Our bounds are the
first to give a scaling for the minimax risk in terms of the number of sensors
used.

@@

Hence, we believe that this paper is the first to give lower bounds on the
minimax error in estimating the location of a point source, which is observed
by sensors through a diffusive medium, and corrupted by additive white Gaussian
noise. We are also the first to analyze the implications physical sensor models
have on such lower bounds. A simplistic theoretical model might assume that the
number of sensors can increase unbounded, while maintaining a fixed SNR at each
sensor. We show that this is not physically possible in our setting, and derive
numerical bounds which have radically different asymptotic behaviour.

@@

In the ``integrator sensor'' model, sensors have a \emph{width} which decreases
as their number grows to infinity. Concretely, the \mbox{$k$-th} sensor which
is located at $s = kS/m$ \emph{occupies} the space $((k{-}\frac{1}{2})S/m,
(k{+}\frac{1}{2})S/m)$.  The signal sensed by this sensor is given by the
$k$-th element of $\v x(\theta)$,
\begin{equation} \label{eq:samples-int-model}
	x_k(\theta) = \int_{(k-\frac{1}{2})S/m}^{(k+\frac{1}{2})S/m} x(s;\theta) ds.
\end{equation}
Here, a white noise process $\epsilon(s)$ additively corrupts the
continuous-space signal $x(s;\theta)$ prior to sampling (see
Fig.~\ref{fig:signal-model}). $\v \epsilon$ is generated by integrating
$\epsilon(s)$ over each sensor's width. This leads to i.i.d.\ noise values at
sensors (since sensors have no overlap), with variance inversely proportional
to $m$, i.e., $\sigma(m) \propto 1/\sqrt{m}$ (refer
Appendix~\ref{app:integrator-snr} in the full version of this
paper~\cite{FullVersion} for an explanation).

@@

	section~\ref{sec:source-localization} under the point sensor model, the
	minimax risk in estimating the location of a point source is lower bounded
	by

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=2.8in]{lb-vs-m-annotated}
	\caption{A numerically generated plot of the lower bound on minimax risk of
		source localization, against the number of uniformly distributed
		sensors, $m$, used to estimate the location (code available
		online~\cite{Code}). The different curves are for different numbers of
		``trials'', $n$ (number of i.i.d.\ sensor observations). The bound
		saturates for large $m$, indicating that for a \emph{fixed} number of
		trials, there is decreasing benefit to using more sensors.  However,
		the plot suggests that with more trials, the bound saturates at
		increasingly larger numbers of sensors. Thus, there may be benefit to
		increasing the number of sensors when you have more trials.}
	\label{fig:numerical}
\end{figure}

Plotting this numerically-computed bound for different numbers of trials
reveals some interesting trends (see Fig.~\ref{fig:numerical}). First, the
bound saturates as $m$ grows large, matching the behaviour of the
information-theoretic bound. Second, the lower bound on error is larger for
smaller numbers of sensors, as we expect. Third and most importantly, as the
number of trials increases, the bound begins to saturate only at a larger
number of sensors. In other words, there is \emph{increased} benefit in using
more sensors, when we \emph{also} have more trials. This substantiates the
intuition given in~\cite{Grover2016Information}, which makes this point using a
Nyquist-rate argument: a larger number of sensors will sample a larger number
of high-frequency components of the source signal (and hence allow us to locate
it better), but only if the noise floor is sufficiently low. Noise variance, in
turn, can be reduced by averaging over multiple trials.

@@

We also note that the minimax bounds from compressive sensing
literature~\cite{AriasCastro2013Fundamental} \emph{do not} apply in our
setting, because those consider the problem of recovering the \emph{whole}
source signal, $f(s;\theta)$. The loss function used there is hence the energy
of the difference signal, $\int \abs{f(s) - \hat f(s)}^2 ds$. We, on the other
hand, are interested in bounding the minimax error in \emph{location}, i.e.,
$\abs{\theta - \hat\theta}^2$.
